Species motif participation provides unique information about species risk of extinction

### 2024-02-16 manuscript/JAE_resub1.tex
tiny revisions

\clearpage
\begin{spacing}{1.0}
% \linenumbers

@@

    A single extinction can, due to the mutual dependencies among species in a food web,  trigger a cascade of additional (secondary) extinctions. 
    A species' risk of secondary extinction depends on its position within a food web~\citep{Santos2021,dunne2009cascading, Eklof2006}.
    For example, higher trophic levels (long paths to basal resources) and lower in-degree (few prey) increase extinction risk~\citep{Allesina2012,binzer2011susceptibility,Eklof2006}.
    These species are more vulnerable because they are impacted by the loss of any species along a long path from basal resources or because they have few alternative resources to replace lost prey.
    The relationships between these relatively simple measures of a species' position in a food web suggest that more detailed measures, such as a species' participation in \emph{meso-scale motifs} are also related to its persistence~\citep{Cirtwill2022Oikos}.

    Motifs are unique patterns of interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale network structure~\citep{Stouffer2007,Stouffer2012}.
    Some motifs are over-represented in stable food webs, suggesting that the more frequently species participate in these motifs, the more stable the network will be~\citep{Borrelli2015a}. 
    The vector of frequencies with which a species appears in each motif --its \emph{motif participation}-- describes the species' position in the meso-scale structure of a food web and, like degree and trophic level, is easily calculated once the structure of the network is known.
    Importantly, motif participation incorporates information on the positions of focal species' interaction partners.
    Two species might have the same trophic level and in-degree but quite different motif participation if one species's prey are strict specialists and the others' prey are generalists (Fig.~\ref{fig:concept}).
    Since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
    This extra information on the focal species' interaction partners may thus provide extra insight when predicting extinction risk.

>>>>

\clearpage
\begin{spacing}{2.0}
\linenumbers

@@

    A single extinction can, due to the mutual dependencies among species in a food web,  trigger a cascade of additional (secondary) extinctions. 
    A species' risk of secondary extinction depends on its position within a food web~\citep{Santos2021,dunne2009cascading, Eklof2006}.
    For example, higher trophic levels (long paths to basal resources) and lower in-degree (few prey) increase extinction risk~\citep{Allesina2012,binzer2011susceptibility,Eklof2006}.
    These species are more vulnerable because they are impacted by the loss of any species along a long path from basal resources or because they have few alternative resources to replace lost prey.
    The relationships between these relatively simple measures of a species' position in a food web suggest that more detailed measures, such as a species' participation in \emph{motifs} are also related to its persistence~\citep{Cirtwill2022Oikos}.

    Motifs are unique patterns of interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale (i.e., intermediate between overall and fine-scale) network structure~\citep{Stouffer2007,Stouffer2012}.
    Some motifs are over-represented in stable food webs, suggesting that the more frequently species participate in these motifs, the more stable the network will be~\citep{Borrelli2015a}. 
    The vector of frequencies with which a species appears in each motif --its \emph{motif participation}-- describes the species' position in the meso-scale structure of a food web and, like degree and trophic level, is easily calculated once the structure of the network is known.
    Importantly, motif participation incorporates information on the positions of focal species' interaction partners.
    Two species might have the same trophic level and in-degree but quite different motif participation if one species's prey are strict specialists and the others' prey are generalists (Fig.~\ref{fig:concept}).
    Since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
    This extra information on the focal species' interaction partners may thus provide extra insight when predicting extinction risk.
